---
title: "GEOG 28602: Final Project"
author: "Josea Evan"
date: "6/1/2021"
output: html_document
runtime: shiny 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(tmap)
library(units)
library(tidyverse)
library(dplyr)
library(leaflet)
library(ggplot2)
library(tidyr)
library(rgdal)
```

## Setting up the Files

Some of our files are in SHP, and some of them are in KMZ. What we want is to standardize everything to SHP files. 

Due to the nature of our `gdal` installation, we are unable to read .kmz files, according to https://mitchellgritts.com/posts/load-kml-and-kmz-files-into-r/. We utilize https://mygeodata.cloud (an open source platform) to help us convert the data from kmz to shp for easier crosswalking calculations. 

Let us plot fist load the files into our project:
```{r }
eld <- st_read("data/elections/eld.shp")
planning <- st_read("data/planning/MP14_PLNG_AREA_WEB_PL.shp") 
```

Setting CRS:
```{r}
planning <- st_transform(planning, st_crs(eld))
```


We can just map out to visually see the maps. First, the election bounds map: 
```{r}
tmap_mode("plot")
map_elections = tm_shape(eld) + tm_fill() + tm_borders() + tm_polygons("Name", legend.show= F) + tm_text("Name", size = 1/3)
map_elections + tm_layout(title = "General Election Bounds (2020)")
```

```{r}
map_planning = tm_shape(planning) + tm_fill() + tm_borders() + tm_polygons("PLN_AREA_N", legend.show= F) + tm_text("PLN_AREA_N", size = 1/3)
map_planning + tm_layout(title = "Planning Areas (2015)")
```

```{r}

#Some formatting for overlaying 2 maps

sixtysix_colors <- list(
  red="#D0445E",
  blue="#0077E0",
  purple="#C92EC4",
  green="#009871",
  orange="#9C7200",
  grey="#009871",
  light_red="#EE6178"
)

theme_map_sixtysix <- function(){ 
  ggthemes::theme_map() %+replace%
    theme(
      text = ggthemes::theme_fivethirtyeight()$text,
      title = ggthemes::theme_fivethirtyeight()$plot.title,
      panel.grid.major = element_line(color="grey90")
    )
}

theme_map_sixtysix <- function(){ 
  ggthemes::theme_map() %+replace%
    theme(
      text = ggthemes::theme_fivethirtyeight()$text,
      title = ggthemes::theme_fivethirtyeight()$plot.title,
      panel.grid.major = element_line(color="grey90")
    )
}
```

Mapping the two on top of each other: 

```{r}
#Creatihng overlapping maps

ggplot(
  eld
  ) +
    geom_sf(
    color = sixtysix_colors$light_red,
    size=1, 
    fill=NA
  ) + 
  geom_text(
    aes(x=X, y=Y, label = ED_CODE),
    data = with(
      eld,
      data.frame(ED_CODE, st_centroid(geometry) %>% st_coordinates())
    ),
    color=sixtysix_colors$light_red,
    size=4,
    fontface="bold"
  ) +
  geom_sf(
    data=planning,
    color = "black",
    size=1,
    fill=NA
  ) +
  geom_text(
    aes(x=X, y=Y, label = PLN_AREA_C),
    data = with(
      planning,
      data.frame(PLN_AREA_C, st_centroid(geometry) %>% st_coordinates())
    ),
    color="black",
    size=4
  ) +
  scale_color_identity(guide=FALSE) +
  theme_map_sixtysix() +
  ggtitle("Election Boundaries (Red) vs \n Planning Areas (Black)")
```

Adopt Areal Weighting Methodology to construct crosswalks: 
We want to obtain a data frame (our crosswalk) that tells us how many percent of the Planning Area data can be attributed to the electoral divisions. 

```{r}
areal_weights <- st_intersection(
  st_make_valid(planning), 
  st_make_valid(eld)
  ) %>%
  mutate(area = st_area(geometry)) %>%
  as.data.frame() %>%
  select(PLN_AREA_N, Name, area) %>% 
  group_by(PLN_AREA_N) %>%
  mutate(prop_of_pln = as.numeric(area / sum(area))
  )
```
Let us do some exploration first and see what the crosswalk "means". 
```{r}
areal_weights %>% filter(Name == "HOLLAND-BUKIT TIMAH")
```
Here we see that to obtain data for the Holland-Bukit Timah constitutency (the electoral boundary I personally belong to), I would take 0.02552 of Bukit Batok's data, 0.81 of Bukit Panjang's, etc. We note that there are also some values in `prop_of_pln` that are very close to zero. These are areas that are probably very trivial (i.e. a very very small intersection) but we will still count them in our final data output for the sake of completeness. 

```{r}
crosswalk = 
  areal_weights %>%
  rename(eld_bounds = Name) %>%
  rename(pln_area = PLN_AREA_N)

write.csv(crosswalk, file="crosswalk.csv")
#ELD bounds = electoral bounds
```


## Using the Crosswalk

To use the crosswalk, we first have to load in a relevant dataset. What we will do is to load in a sample dataset, use the crosswalk once, and then produce a function that will allow us to replicate the calculations by just passing a relevant dataset. 

### Input Income Dataset

```{r}
income_pln <- read_csv("data/raw/income.csv")
```


### Cleaning and Transforming the Income Dataset
```{r}

# Let's just keep the columns that we want and throw out the rest.
clean_inc =
income_pln %>%
  select(level_1, level_3, value) %>% 
  rename(breakdown = level_1) %>%
  rename(pln_area = level_3)

joined_income = 
  left_join(clean_inc, crosswalk, by = "pln_area") %>% 
  mutate(scaled = as.numeric(value * prop_of_pln))

```
Using the joined dataset, we group by breakdown (which is the "income level") and electoral bounds. We then sum up the "scaled" values per intersection (e.g. Planning Area 1 intersects with Election Bound A, Planning Area 2 intersects with Election Bound A), and then group by electoral bounds.
```{r}
test_income = 
  joined_income %>% 
  group_by(breakdown, eld_bounds) %>%
  mutate(sum_eld = sum(scaled)) %>%
  arrange(breakdown, eld_bounds)

test_income 
```

To better understand this, let's just filter out a small proportion of the dataset to see what is going on.

```{r}
test_income %>% filter(eld_bounds == "HOLLAND-BUKIT TIMAH", breakdown == "$2,000 - $2,999")
```
This dataset tells us that within the Holland-Bukit Timah Constituency, the 'value' of those earning between $2,000 - $2,999 of income is 13.60. We will go into detail what these "values" mean later. 

### Interpreting Values 

Values in this case represent the number of people [Resident Working Persons Aged 15 Years and Over by Planning Area and Gross Monthly Income from Work, 2015], in thousands. What we want to do is to convert each value to a percentage of that election bounds's population. 

First, we need to calculate the electoral bounds's total population (our denominator, if you will). 
```{r}
test_income %>% 
  filter(breakdown == "Total")
```
To calculate Aljunied's total population, we will 'search up' Aljunied in `eld_bounds`, sum up the corresponding values in `sum_eld`, and then divide it by the number of rows in Aljunied. We can write a for loop to do that:

```{r}
total_income =
test_income %>% 
  filter(breakdown == "Total") %>%
  group_by(eld_bounds) %>% 
  summarise(pop_in_000 = mean(sum_eld)) ## Mean is the same as adding and then dividing!

total_income
```

We can write a function that extracts the correct population value for that level. 

```{r}
## List of Levels
inc_levels <- list('Below $1,000', 
                   '$1,000 - $1,999',
                   '$2,000 - $2,999', 
                   '$3,000 - $3,999', 
                   '$4,000 - $4,999',
                   '$5,000 - $5,999', 
                   '$6,000 - $6,999', 
                   '$7,000 - $7,999', 
                   '$8,000 - $8,999',
                   '$9,000 - $9,999', 
                   '$10,000 - $10,999', 
                   '$11,000 - $11,999', 
                   '$12,000 & Over'
                   )
```

```{r}
#Calculate each level of income as a percentage of total population
income_value_pct <- function(level) {
  test_income %>%
    filter(breakdown == level) %>%
    group_by(eld_bounds) %>% 
    summarise(pop_in_000 = mean(sum_eld)) %>% 
    mutate(pct = 100* (pop_in_000 / total_income$pop_in_000)) %>%
    mutate(inc_level = level) # We include this so we know what we are looking at 
}
```

```{r}
eld_income <- lapply(inc_levels, income_value_pct) %>% bind_rows() %>% drop_na()
eld_income
```

### Visualizing Income Data

Now that we have the consolidated income data by electoral bounds in the `eld_income` dataframe, we can begin some visualizations of the data. 

First, let us generate some histograms. 

To generate histograms, we need to "delimit" our eld_income dataset firstly by bounds. 


```{r, fig.width=15, fig.height=15}

ggplot(eld_income, aes(fill=factor(inc_level, levels=rev(inc_levels)), y=pct, x=eld_bounds)) + 
  geom_bar(position="fill", stat="identity") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size=20),
        axis.title.x = element_text(size=30),
        axis.title.y = element_text(size = 30),
        axis.text.y = element_text(size=20),
        legend.box.background = element_rect(),
        legend.box.margin = margin(6, 6, 6, 6), 
        legend.key.size = unit(2,'cm'), 
        legend.title = element_text(size=18, face='bold'),
        legend.text = element_text(size=15)) + xlab("Electoral Boundary Name") +
  ylab("Percentage of Population") + labs(fill = "Income Levels") + coord_flip()
```

### Input Education Dataset
We do the same with the education dataset. 
```{r}
edu_pln <- read.csv("data/raw/edu_pln.csv")
```

```{r}

# Let's just keep the columns that we want and throw out the rest.
clean_edu =
edu_pln %>%
  select(level_1, level_3, value) %>% 
  rename(breakdown = level_1) %>%
  rename(pln_area = level_3)

joined_edu = 
  left_join(clean_edu, crosswalk, by = "pln_area") %>% 
  mutate(scaled = as.numeric(value * prop_of_pln))

```

```{r}
test_edu = 
  joined_edu %>% 
  group_by(breakdown, eld_bounds) %>%
  mutate(sum_eld = sum(scaled)) %>%
  arrange(breakdown, eld_bounds)

test_edu 
```


### Interpreting Values 

Values in this case represent the number of people in thousands and their respective highest education attained. We convert this to a percentage value of total population.  

First, we need to calculate the electoral bounds's total population (our denominator, if you will). 

```{r}
total_edu =
test_edu %>% 
  filter(breakdown == "Total") %>%
  group_by(eld_bounds) %>% 
  summarise(pop_in_000 = mean(sum_eld)) ## Mean is the same as adding and then dividing!

total_edu
```

We can write a function that extracts the correct population value for that level. 

```{r}
## List of Levels
edu_levels <- list('No Qualification', 
                   'Primary',
                   'Lower Secondary', 
                   'Secondary', 
                   'Post-Secondary (Non-Tertiary)',
                   'Polytechnic', 
                   'Professional Qualification and Other Diploma', 
                   'University'
                   )
```

```{r}
#Calculate each level of edu as a percentage of total population
edu_value_pct <- function(level) {
  test_edu %>%
    filter(breakdown == level) %>%
    group_by(eld_bounds) %>% 
    summarise(pop_in_000 = mean(sum_eld)) %>% 
    mutate(pct = 100* (pop_in_000 / total_edu$pop_in_000)) %>%
    mutate(edu_level = level) # We include this so we know what we are looking at 
}
```

```{r}
eld_edu <- lapply(edu_levels, edu_value_pct) %>% bind_rows() %>% drop_na()
eld_edu
```

### Visualizing Education Data

```{r, fig.width=15, fig.height=15}
#Make education barplot
ggplot(eld_edu, aes(fill=factor(edu_level, levels=rev(edu_levels)), y=pct, x=eld_bounds)) + 
  geom_bar(position="fill", stat="identity") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size=20),
        axis.title.x = element_text(size=30),
        axis.title.y = element_text(size = 30),
        axis.text.y = element_text(size=20),
        legend.box.background = element_rect(),
        legend.box.margin = margin(6, 6, 6, 6), 
        legend.key.size = unit(2,'cm'), 
        legend.title = element_text(size=18, face='bold'),
        legend.text = element_text(size=15)) + xlab("Electoral Boundary Name") +
  ylab("Percentage of Population") + labs(fill = "Education Levels") + coord_flip()

```

We also import voting data. We express voting data in the column `vote_pap_pct`, which shows the percentage of the population that voted for the People's Action Party, who is the ruling party that contests in every single constitutency (electoral area). 
```{r}
voting <- read_csv("data/raw/voting.csv")
voted_pap = voting %>% filter(year==2020, party=="PAP") %>% mutate(vote_pap_pct = as.numeric(vote_percentage) * 100) %>% select(constituency, vote_pap_pct) # Data from the 2020 Elections
```


Join numeric data to the main dataset of electoral data. 

```{r}

# We pivot the data to get the income and ed
pivoted_income = eld_income %>% select(-"pop_in_000") %>% pivot_wider(names_from = inc_level, values_from = pct) 
pivoted_edu = eld_edu %>% select(-"pop_in_000") %>% pivot_wider(names_from = edu_level, values_from = pct)

eld_2 = eld %>% select(c("Name", "geometry")) # We make a copy of the Electoral Boundary file with the columns we want

join_income_eld = inner_join(eld_2, pivoted_income, by=c("Name"='eld_bounds')) # We join income data to eld data first
join_edu_data = inner_join(join_income_eld, pivoted_edu, by=c("Name"="eld_bounds")) # Join Edu Data with income Data
eld_data = 
  inner_join(join_edu_data, voted_pap, by=c("Name"="constituency")) %>%
  mutate(across(where(is.numeric), round, 2)) #Round values

# tibble(eld_data)
# write.csv(eld_data, file="eld_data.csv")
# write.csv(eld_income, file="eld_income.csv")
# write_sf(eld_data, "eld_data.shp")

```

Now that we have a completed and cleaned datasaet at the electoral boundary level, we can begin to create some cartographical visuatlizations.

Make Maps 

First, we generate maps that show the percentage of the population that voted for the ruling People's Action Party. Before we do that, let us make some custom color palettes to visualize our data. 

```{r}
mypal <- c('#edf8fb', '#b3cde3', '#8c96c6','#8856a7','#810f7c')
tmap_mode("plot")
tm_shape(eld_data)  + tm_polygons("vote_pap_pct", title = "Percentage Voted for PAP in 2020 Elections", palette = mypal) + tm_borders() +  tm_text("Name", size = 1/3)
```

```{r}
tmap_mode("view")
income_map = 
  tm_shape(eld_data) + tm_text("Name", size = 1/1.3) + tm_borders(lwd = 2) + 
  tm_polygons("vote_pap_pct", title = "Percentage Voted for PAP",
              popup.vars=c(
                "Income Below $1,000: " = "Below $1,000", 
                "Income $1,000 - $1,999: " = "$1,000 - $1,999",
                "Income $2,000 - $2,999: " = "$2,000 - $2,999",
                "Income $3,000 - $3,999: " = "$3,000 - $3,999",
                "Income $4,000 - $4,999: " = "$4,000 - $4,999",
                "Income $5,000 - $5,999: " = "$5,000 - $5,999",
                "Income $6,000 - $6,999: " = "$6,000 - $6,999",
                "Income $7,000 - $7,999: " = "$7,000 - $7,999",
                "Income $8,000 - $8,999: " = "$8,000 - $8,999",
                "Income $9,000 - $9,999: " = "$9,000 - $9,999",
                "Income $10,000 - $10,999: " = "$10,000 - $10,999",
                "Income $11,000 - $11,999: " = "$11,000 - $11,999",
                "Income above $12,000: " = "$12,000 & Over", 
                "Percentage who voted for PAP: " = "vote_pap_pct")
                ) + tm_layout(title="Income (2015) and PAP Vote Share (2020)")

tmap_leaflet(income_map, show = TRUE, add.titles=TRUE)
```

```{r}

edu_map = 
  tm_shape(eld_data) + tm_text("Name", size = 1/1.3) + tm_borders(lwd = 2) + 
  tm_polygons("vote_pap_pct", title = "Percentage Voted for PAP",
              popup.vars=c(
                "No Qualification: " = "No Qualification", 
                "Primary: " = "Primary", 
                "Lower Secondary: " ="Secondary", 
                "Secondary: " = "Secondary",
                "Post-Secondary (Non-Tertiary): " = "Post-Secondary (Non-Tertiary)",
                "Polytechnic: " = "Polytechnic",
                "Professional Qualification and Other Diploma: " = "Professional Qualification and Other Diploma",
                "University: " = "University", 
                "Percentage who voted for PAP: " = "vote_pap_pct")
                ) + tm_layout(title="Highest Edu Qualification (2015) and PAP Vote Share (2020)")

tmap_leaflet(edu_map, show = TRUE, add.titles=TRUE)

```

We take the visualizations a step further by creating an R Shiny app that allows us to view the data in a centralized fashion. The R Shiny App should allow us to compare data in an interactive manner at the electoral boundary level between electoral boundaries.

```{r, echo=FALSE}
shinyApp(

  ui = fluidPage(
  
  # App title ----
  titlePanel("Singapore Electoral Boundary Data Explorer"),
  
  # Sidebar layout with input and output definitions ----
  sidebarLayout(
    sidebarPanel(
      h3("Make your choice"),
      helpText("Explore Singapore data based on selected variables"),
      selectInput(
      "selectedCol1",
      "Select 1st boundary to compare",
      choices = list("ALJUNIED",
                     "ANG MO KIO",
                     "BISHAN-TOA PAYOH",
                     "BUKIT BATOK",
                     "BUKIT PANJANG",
                     "CHUA CHU KANG",
                     "EAST COAST",
                     "HOLLAND-BUKIT TIMAH",
                     "HONG KAH NORTH",
                     "HOUGANG",
                     "JALAN BESAR",
                     "JURONG",
                     "KEBUN BARU",
                     "MACPHERSON",
                     "MARINE PARADE",
                     "MARSILING-YEW TEE",
                     "MARYMOUNT",
                     "MOUNTBATTEN",
                     "NEE SOON",
                     "PASIR RIS-PUNGGOL",
                     "PIONEER",
                     "POTONG PASIR",
                     "PUNGGOL WEST",
                     "RADIN MAS",
                     "SEMBAWANG",
                     "SENGKANG",
                     "TAMPINES",
                     "TANJONG PAGAR",
                     "WEST COAST",
                     "YIO CHU KANG",
                     "YUHUA"),
      selected = "Bound 1"
    ),
    selectInput(
      "selectedCol2",
      "Select 2st boundary to compare",
      choices = list("ALJUNIED",
                     "ANG MO KIO",
                     "BISHAN-TOA PAYOH",
                     "BUKIT BATOK",
                     "BUKIT PANJANG",
                     "CHUA CHU KANG",
                     "EAST COAST",
                     "HOLLAND-BUKIT TIMAH",
                     "HONG KAH NORTH",
                     "HOUGANG",
                     "JALAN BESAR",
                     "JURONG",
                     "KEBUN BARU",
                     "MACPHERSON",
                     "MARINE PARADE",
                     "MARSILING-YEW TEE",
                     "MARYMOUNT",
                     "MOUNTBATTEN",
                     "NEE SOON",
                     "PASIR RIS-PUNGGOL",
                     "PIONEER",
                     "POTONG PASIR",
                     "PUNGGOL WEST",
                     "RADIN MAS",
                     "SEMBAWANG",
                     "SENGKANG",
                     "TAMPINES",
                     "TANJONG PAGAR",
                     "WEST COAST",
                     "YIO CHU KANG",
                     "YUHUA"),
      selected = "Bound 2"
    ),
    selectInput("Map_Choice", 
                label = "View electoral boundary level data on an interactive map of Singapore",
                choices = list("Income Data", 
                               "Education Data"),
                selected = "Income Data"),
    
    h4("About"),
    p("This project was completed by Josea Evan, as part of coursework for GIS III at the University of Chicago. Contact: josea [at] uchicago.edu"),
    h4("Data Source"),
    p("This data is from the Singapore Census. Electoral Data from 2020 Singapore General Election. Income and Education Data crosswalked from 2015 Household Survey.")
  ),
    
    # Main panel for displaying outputs ----
    mainPanel(
      # Output: Tabset w/ plot, summary, and table ----
      tabsetPanel(type = "tabs",
                  tabPanel("Education Data", plotOutput("distPlot")),
                  tabPanel("Income Data", plotOutput("incPlot")),
                  tabPanel("Map", leafletOutput("working_map"))

      )
      
    )
  )
),

  server <- function(input, output) {
  
  output$distPlot <- renderPlot({
    test_edu = 
      eld_edu %>% filter(eld_bounds %in% c(input$selectedCol1, input$selectedCol2))
    
    
    ggplot(test_edu, aes(fill=factor(edu_level, levels=rev(edu_levels)), y=pct, x= "Boundary Name")) + 
      geom_bar(position="fill", stat="identity") + facet_grid(~eld_bounds) + 
      theme(legend.box.background = element_rect(),
            legend.title = element_text(face='bold')) + 
      labs(fill = "Highest Education Level Attained (2015)") +
      ylab("Percentage of Population") 
    
  })

  #Generate a Violin Plot of Variables 
  output$incPlot <- renderPlot({
    test_inc = 
      eld_income %>% filter(eld_bounds %in% c(input$selectedCol1, input$selectedCol2))
    
    
    ggplot(test_inc, aes(fill=factor(inc_level, levels=rev(inc_levels)), y=pct, x= "Boundary Name")) + 
      geom_bar(position="fill", stat="identity") + facet_grid(~eld_bounds) + 
      theme(legend.box.background = element_rect(),
            legend.title = element_text(face='bold')) + 
      labs(fill = "Gross Monthly Income from Work (2015)") +
      ylab("Percentage of Population") 
  })
  
  output$working_map <- renderLeaflet({
    
    map <- switch(input$Map_Choice, 
                  "Income Data" = income_map,
                  "Education Data" = edu_map)
    
    working_map <- tmap_leaflet(map)
    
  })
},

  options = list(height = 1000)
)
```
